---
title: "Pressure Volume Curves"
author: "German Vargas G."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pressure volume curves:

These are measurements commonly done by plant ecophysiologists in order to obtain the turgor loss point (TLP, MPa) of leaves (a.k.a wilting point). 

This is a tutorial on how to analyze pressure volume curves data. For more details on  how to perform pressure volume curves using a pressure chamber see: <http://prometheuswiki.org/tiki-index.php?page=Leaf+pressure-volume+curve+parameters>.

The functions provided here are based on the method provided here:

- Prometheus wiki: <http://prometheuswiki.org/tiki-index.php?page=Leaf+pressure-volume+curve+parameters>

- Koide, R. T., Robichaux, R. H., Morse, S. R., & Smith, C. M. (1989). Plant water status, hydraulic resistance and capacitance. In _Plant physiological ecology_ (pp. 161-183). Springer, Dordrecht. <https://link.springer.com/chapter/10.1007/978-94-009-2221-1_9>

- Sack, L., Cowan, P. D., Jaikumar, N., & Holbrook, N. M. (2003). The ‘hydrology’of leaves: co‐ordination of structure and function in temperate woody species. _Plant, Cell & Environment_, 26(8), 1343-1356. <https://onlinelibrary.wiley.com/doi/full/10.1046/j.0016-8025.2003.01058.x>

- Tyree, M. T., & Hammel, H. T. (1972). The measurement of the turgor pressure and the water relations of plants by the pressure-bomb technique. Journal of Experimental Botany, 23(1), 267-282.<https://academic.oup.com/jxb/article-abstract/23/1/267/623193>

## Load the functions
```{r}
source(file = "scripts/functions/water_mass.R")#directory to the scrpt files containing the functions
source(file = "scripts/functions/pvc_fitting_v4.R")#directory to the script files containing the functions
```

## Load the data
```{r}
trial <- read.csv(file = "data/trial_PVcurve.csv")
trial
```

## Fiting the PV curves

### No leaf area provided
 
```{r}
pvc_fitting(total_mass = fresh_mass_g,# fresh mass during the measurements
            dry_mass = dry_mass_g,# leaf dry weight
            extra_mass = extra_mass_g,# additional mass as rubber gaskets, whirlpac bags, etc...
            water_pot = water_pot_Mpa,#water potential during the measurements
            ip =1,# initial point, in case the sample is overhydrated
            bkp = 6,# break point of the curve 
            leafA = FALSE,# leaf area provided yes or no
            data = trial, ID="Eugenia procera")# dataset with the curve, just need water potentials and weights
```


### Leaf area provided

In this case the output will contain absolute capacitance at full turgor

```{r}
pvc_fitting(total_mass = fresh_mass_g,# fresh mass during the measurements
            dry_mass = dry_mass_g,# leaf dry weight
            extra_mass = extra_mass_g,# additional mass as rubber gaskets, whirlpac bags, etc...
            water_pot = water_pot_Mpa,#water potential during the measurements
            ip =1,# initial point, in case the sample is overhydrated
            bkp = 6,# break point of the curve 
            leafA = T,# leaf area provided yes or no
            LA = trial$leaf_area_cm2[1],#columen with leaf area
            data = trial, ID="Eugenia procera")#
```

